---
title: DANL Project
subtitle: "hoopR: The SportsDataverse's R Package for Men's Basketball Data."
author: 
  - Name Tristan Costello
  - Name Quinn Campbell
  - Name Joshua Lefkowicz
  - Name Andrew Thompson
  - Name Hannah Chisholm
toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true
fig-width: 9
execute:
  code-tools: true
  message: false
  warning: false
from: markdown+emoji
editor: 
  markdown: 
    wrap: 72
---

```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```

# Introduction

About this project :clap:

This project seeks to examine and highlight the disparities in team
performance between their home court and away games. Through a detailed
analysis, we aim to uncover insights into how external factors, such as
the playing venue, influence sporting outcomes. By exploring these
nuances, we intend to provide valuable perspectives on the intricate
dynamics that distinguish home and away performances in competitive
sports.

# Data

The dataset nba_pbp_ny comprises a subset of the NBA play-by-play data,
specifically selected based on a criteria related to this years games.
In the context of the project, this subset includes only those teams or
players that consistently participated and demonstrated a sustained
level of activity in the NBA across 2023-2024 season thus far. This
criterion was utilized as a proxy for gauging the enduring relevance or
popularity of specific teams or players during that period. The aim is
to investigate and compare the performance metrics of these consistently
active entities, both at their home court and away, providing insights
into how their sustained presence correlates with their gameplay
dynamics over the specified timeframe. :basketball:

## Summary Statistics

```{r}
nba_pbp_ny <- read_csv("https://tristancostello.github.io/nba_hoopr_ny.csv")
```

```{r}
#| eval: false
rmarkdown::paged_table(nba_pbp_ny) 
```

```{r}
skim(nba_pbp_ny) %>% 
  select(-n_missing)
```

## Home Score and Team

The following boxplot shows the home score (`home_score`) of the home
team (`home_team_name`) :basketball

```{r}
ggplot(data = 
         nba_pbp_ny %>% 
         filter(home_team_name %in% c("Utah", "New York","Boston"))) +
  geom_boxplot(aes(x = home_team_name, y = home_score, 
                   fill = home_team_name),
               show.legend = FALSE)
```

## Away Score and Team

The following boxplot shows the away score (`away_score`) of the away
team (`away_team_name`) :basketball

```{r}
ggplot(data = 
         nba_pbp_ny %>% 
         filter(away_team_name %in% c("Utah", "New York","Boston"))) +
  geom_boxplot(aes(x = away_team_name, y = away_score, 
                   fill = away_team_name),
               show.legend = FALSE)
```

## Home Score and Team

The following line graph shows the home score (`home_score`) of the home
team (`home_team_name`) :basketball

```{r}
nba_pbp_ny %>%
  filter(home_team_name %in% c("Utah", "New York", "Boston")) %>%
  ggplot(aes(x = as.factor(home_team_name), y = home_score, group = home_team_name, color = home_team_name)) +
  geom_line() +
  labs(title = "Home Team Scores Over Time",
       x = "Home Team Name",
       y = "Home Team Score") +
  theme_minimal()
```

## Away Score and Team

The following line graph shows the away score (`away_score`) of the away
team (`away_team_name`) :basketball

```{r}
nba_pbp_ny %>%
  filter(away_team_name %in% c("Utah","New York", "Boston")) %>%
  ggplot(aes(x = as.factor(away_team_name), y = away_score, group = away_team_name, color = away_team_name)) +
  geom_line() +
  labs(title = "Away Team Scores Over Time",
       x = "Away Team Name",
       y = "Away Team Score") +
  theme_minimal()
```

## Why do teams tend to do better at home?

Home court advantage serves as a decisive factor across all professional
sports, providing teams with a significant edge through a positive
atmosphere and heightened motivation fueled by fan support. The familiar
surroundings and enthusiastic cheers contribute to a sense of
camaraderie, inspiring players to deliver standout performances on their
home court. This intricate interplay between psychological factors and
tangible gameplay dynamics underscores the enduring impact of home court
advantage in shaping the outcomes of competitive matches.

## Why do teams score more against other teams?

Teams may score more against certain opponents due to a variety of
factors. These could include differences in team strategies, strengths
and weaknesses, player matchups, and tactical variations that exploit
vulnerabilities in the opposing team's defense. Additionally, factors
such as team morale, motivation, and the specific dynamics of each match
can influence scoring outcomes, creating variations in performance
against different opponents. Understanding the specific elements at
play, both strategically and psychologically, contributes to unraveling
the complexities behind why teams may exhibit higher-scoring
performances against certain adversaries.

## Are there specific players associated with higher scoring games?

Yes, specific basketball players are often associated with
higher-scoring games, and their individual performances can
significantly impact the overall scoring of their teams. Star players
known for their scoring abilities, such as prolific scorers or
sharpshooters, tend to contribute more points to their team's total.
Factors such as a player's skill set, shooting accuracy, scoring
versatility, and ability to create scoring opportunities can all
influence their association with higher-scoring games. Analyzing
individual player performances provides insights into the key
contributors to a team's overall scoring and helps identify players who
consistently excel in putting points on

## Can we identify any variables that correlate strongly with team performance?

Exploring the correlation between various variables and team performance
is crucial for gaining insights into the factors that contribute to a
team's success. Key variables to consider include player statistics
(e.g., points per game, assists, rebounds), team dynamics, coaching
strategies, and game-related metrics (e.g., field goal percentage,
three-point percentage, turnovers). By conducting a comprehensive
analysis, we can identify which of these factors exhibit a strong
correlation with team performance, providing valuable information for
coaches, analysts, and teams seeking to optimize their strategies and
enhance overall on-court success.
